---
title: "Describe models included in rtedem"
author: "K Todd-Brown (ktoddbrown@gmail.com)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}

#@param param.df is a data frame with (modelStr, pool, lower, upper)
plotModelTurnovers <- function(param.df){
  
  param.df$modelStr<- factor(param.df$modelStr)
  param.df$pool <- factor(param.df$pool)
  param.df$x <- unclass(param.df$modelStr)
  #print(param.df)
  axisLabels <- unique(param.df[, c('x', 'modelStr')])
  p <- ggplot(param.df) + geom_rect(aes(ymin=x-0.25, ymax=x+0.25, xmin=lower, xmax=upper, fill=pool)) + 
    geom_point(aes(y=x, x=val)) +
    scale_x_log10() + scale_y_continuous(breaks=axisLabels$x, labels=axisLabels$modelStr) + 
    xlab('Turnover time range[yr]') + ylab('')
  return(p)
}

plotModelTransfers <- function(parList, numPool = 5){
  poolOrder <- lapply(parList, function(x)order(-1*x$true[grepl('K', x$name)]))
  countPools <- unlist(lapply(poolOrder, length))
  
  p <- ggplot(data.frame(x=rep(1:numPool, each=numPool), y=-1*rep(1:numPool, times=numPool))) + 
    geom_rect(aes(xmin=x-0.5, xmax=x+0.5, ymin=y-0.5, ymax=y+0.5), alpha=0, color='black')
  
  modelArr <- names(countPools)[countPools == numPool]
  if(length(modelArr) > 4){
    warning('More then 4 pools match desired pool number; truncating.')
  }
  
  for(modelIndex in 1:4){
    modelStr <- modelArr[modelIndex]
    ii <- as.numeric(gsub('^A', '', parList[[modelStr]]$name[grepl('^A', parList[[modelStr]]$name)]))
    transfers <- parList[[modelStr]]$true[grepl('^A', parList[[modelStr]]$name)]
    fromPool <- poolOrder[[modelStr]][rep(1:numPool, each=numPool)[ii]]
    toPool <- -1*poolOrder[[modelStr]][rep(1:numPool, times=numPool)[ii]]
    voffset <- c(0.25, 0.25, -0.25, -0.25)[modelIndex]
    hoffset <- c(0.25, -0.25, 0.25, -0.25)[modelIndex]
    col <- c('red', 'blue', 'green', 'purple')[modelIndex]
    p <- p + annotate('text', x=fromPool+hoffset, y=toPool+voffset, label=transfers, color=col)
  }
  return(p)
}
```

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
